---
author: "University of Melboune"
date: "`r format(Sys.time(), '%a %Y-%b-%d')`"
output:
  html_document:
    keep_md: yes
    theme: readable
    toc: true
    toc_float: true
    code_folding: hide
  rmdformats::material:
    highlight: kate
params:
  batch_name: "x"
  results_dir: "x"
  title: "Data Validation Summary"
title: "`r paste0(params$batch_name, '\n\n', params$title)`"
---

```{r load-pkgs, message=FALSE}
require(dplyr)
require(readr)
require(glue)
require(DT)
require(rock)
require(devtools)
require(kableExtra)
```

```{r render_interactively, eval=FALSE, echo=FALSE}
render_me <- function() {
  rmarkdown::render(
    "validate_report.Rmd",
    params = list(
      batch_name = "transfer_20180921162830", 
      results_dir = "/Users/pdiakumis/Desktop/projects/umccr/woof/data/aws_validate/results/transfer_20180921162830")
  )
}

render_me()
```

Introduction
------------
This report summarises the results from running different data validation tools
on the BAM, FASTQ and VCF files in ``r params$batch_name``.

Tools used include:

- [fqtools](https://github.com/alastair-droop/fqtools)
- [samtools quickcheck](https://github.com/samtools/samtools)
- [vcf-validator](https://github.com/EBIvariation/vcf-validator)


What was meant to be given
--------------------------
The ``r params$batch_name`` directory contains a `manifest.txt` file, which 
lists the files that are meant to be in the directory:

```{r}
manifest <-
  read_tsv(file.path(params$results_dir, "manifest.txt"), skip = 1, col_types = "ccc",
           col_names = c("id", "data_type", "file")) %>%
  mutate(ftype = case_when(
    data_type == "Mitochondrial" ~ "FASTQ_Mito",
    data_type == "WES (SureSelect CRE)" ~ "FASTQ_Exome",
    data_type == "WES Alignment" ~  "BAM",
    data_type == "WES Variants" ~ "VCF",
    data_type == "WES VCF" ~ "VCF",
    TRUE ~ "OTHER"
  )) %>% 
  arrange(id)

datatable(manifest, extensions = c('Scroller'), 
          options = list(scroller = TRUE, scrollX = TRUE, scrollY = 300))
```

What was actually given
-----------------------
The actual contents of the directory are:

```{r}
given_files <-
  read_delim(file.path(params$results_dir, "file_sizes.txt"), delim = ",", col_types = "cc", 
             col_names = c("size", "file")) %>% 
  mutate(size = readr::parse_number(size)) %>%
  mutate(size = fs::as_fs_bytes(size))

datatable(given_files, extensions = c('Scroller'), 
          options = list(scroller = TRUE, scrollX = TRUE, scrollY = 300))
```


Results
-------
Let's check if we got what we meant to:

```{r}
sample_ids <- readxl::read_xlsx(
  "../../data/aws_validate/sample_ids/DEIDENTIFIED_Mito_Data_Extraction_WES_VCGS_180815.xlsx") %>% 
  dplyr::select(study_num = `Study Number:`, 
                sex = `Sex of patient:`, 
                id = `Clinical Exome Sample Number`) %>% 
  dplyr::slice(-n())

comp_df <- dplyr::left_join(manifest, given_files, by = "file") %>% 
  dplyr::left_join(sample_ids, by = "id")

all(c(nrow(manifest), nrow(given_files)) == nrow(comp_df)) &&
  all(complete.cases(comp_df))

```

Now let's see if we got the same type of files for all samples:

```{r}
addmargins(table(comp_df$ftype, useNA = "ifany")) %>% 
  dplyr::as_tibble() %>% 
  purrr::set_names(c("File_Type", "n")) %>% 
  knitr::kable() %>% 
  kableExtra::kable_styling(full_width = FALSE) %>% 
  kableExtra::column_spec(1, bold = TRUE)

f_per_sample <- comp_df %>%
  dplyr::group_by(study_num, ftype) %>% 
  dplyr::summarise(n = n()) %>% 
  tidyr::spread(ftype, n) %>% 
  tidyr::replace_na(list(BAM = 0, FASTQ_Exome = 0, FASTQ_Mito = 0, VCF = 0)) %>% 
  dplyr::arrange(study_num)


f_per_sample %>%  
  datatable(extensions = c('Scroller'), 
            options = list(scroller = TRUE, scrollX = TRUE, scrollY = 300)) %>% 
  formatStyle(
    2:5,
    backgroundColor = styleInterval(0, c('#e6c3c3', '#c3e6c3'))
  )
  
```

Several samples are missing Mitochondrial FASTQ files, but the Exome FASTQs, BAMs
and VCFs are available for all samples.

Now let's see if the datasets are valid.

* `fqtools` was run on the FASTQ and BAM files
* `samtools quickcheck` was run on the BAM files
* `vcf-validator` was run on the VCF files
* `md5sum` was run on all files

```{r}
fqtools_fnames <- list.files(
  file.path(params$results_dir, "validate/fqtools"), 
  full.names = TRUE)
fq_res <- purrr::map(fqtools_fnames, read_lines, n_max = 1) %>%
  purrr::set_names(sub("_valid_summary.txt", "", basename(fqtools_fnames))) %>% 
  dplyr::bind_rows() %>% 
  tidyr::gather(key = "file_name", value = "fqtools")

samquick_fnames <- list.files(file.path(params$results_dir, "validate/samtools_quickcheck"), full.names = TRUE)
samquick_res <- purrr::map(samquick_fnames, read_lines, n_max = 1) %>%
  purrr::set_names(sub("_valid_summary.txt", "", basename(samquick_fnames))) %>% 
  dplyr::bind_rows() %>% 
  tidyr::gather(key = "file_name", value = "quickcheck") %>% 
  dplyr::mutate(quickcheck = toupper(quickcheck))

vcfval_fnames <- list.files(file.path(params$results_dir, "validate/vcfvalidator"), full.names = TRUE)
vcfval_res <- purrr::map(vcfval_fnames, read_lines, n_max = 1) %>%
  purrr::set_names(sub("_valid_summary.txt", "", basename(vcfval_fnames))) %>% 
  purrr::map(function(x) ifelse(length(x) > 0, "ok", "fail")) %>% 
  dplyr::bind_rows() %>% 
  tidyr::gather(key = "file_name", value = "vcfvalidate") %>% 
  dplyr::mutate(vcfvalidate = toupper(vcfvalidate))

md5_fnames <- list.files(file.path(params$results_dir, "validate/md5sum"), full.names = TRUE)
md5_res <- purrr::map(md5_fnames, function(x) { 
  readr::read_table2(x, col_names = c("md5sum", "file_name"), col_types = "cc") %>%
    dplyr::mutate(file_name = basename(file_name)) %>% 
    dplyr::select(file_name, md5sum)
  }) %>%
  purrr::set_names(sub("_md5sum.txt", "", basename(md5_fnames))) %>% 
  dplyr::bind_rows()
```

### All together

```{r}
# start with md5 since it's been run on all files
res <-  dplyr::left_join(md5_res, fq_res, by = "file_name") %>% 
  dplyr::left_join(samquick_res, by = "file_name") %>% 
  dplyr::left_join(vcfval_res, by = "file_name") %>% 
  dplyr::select(file_name, fqtools:vcfvalidate, md5sum)

datatable(res, extensions = c('Scroller'), 
          options = list(scroller = TRUE, scrollX = TRUE, scrollY = 600),
          caption = htmltools::tags$caption(
            style = 'caption-side: bottom; text-align: center;',
            htmltools::em('Green: OK; White: NA; Red: FAIL'))) %>% 
  formatStyle(
    2:4,
    backgroundColor = styleEqual(c(NA, "FAIL", "OK"), c('white', '#e6c3c3', '#c3e6c3'))
  )

```

